---
title: "Motor Trend Car Analysis"
author: "Juan Antonio Morales, juanan4290@gmail.com"
date: "February 15, 2017"
output:
  md_document:
    variant: markdown_github
---

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "figure/"
)
```

Motor Trend Car Analysis
========================

###Summary
In this project we are going to work for *Motor Trend*, a magazine about the automobile industry. They are interested in exploring the relationship between a set of variables and miles per gallon. They are particularly interested in the following two questions:

1. Is an automatic or manual transmission better for MPG?
2. Quantify the MPG difference between automatic and manual transmissions

###Exploratory Data Analysis
The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

```{r data}
library(datasets) 
data(mtcars)
```


Some exploratory data analysis:
```{r exploratory data analysis}
str(mtcars)
```

For the analysis we have to transform the class of some variables variable into a factor:
```{r am class}
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am,labels=c('Automatic','Manual'))
```

###1.Is an automatic or manual transmission better for MPG?
In Appendix 1 we can see a boxplot of MPG by transmission types (Appendix 2). According to Appendix 2 boxplot, it seems that manual transmission is better in miles/(US) gallon than automatic transmission. We can do a t-test to reject the null hypothesis (there is no difference in MPG between manual and automatic transmission).

```{r t-test}
t.test(mpg ~ am,data=mtcars)
```

Based on t-test, **p-value = 0.001374** < 0.05, so we can reject the null hypothesis and hence __manual transmission is better than automatic transmission for MPG.__


###2.Quantify the MPG difference between automatic and manual transmissions
To quantify the MPG difference between automatic and manual transmissions we are going to fit a lineal model between **mpg** and **am** variables.

```{r fit}
fit1=lm(mpg ~ am,data=mtcars)
summary(fit1)
```

From this model, we can say that __automatic run at 17.15 mpg, while manual have 7.24 more mpg__ and about 35.98% of the variance is explained by this model.

Multivariable linear regression model:
```{r fit2}
fit2=lm(mpg ~ am + wt, data=mtcars)
fit3=lm(mpg ~ am + wt + cyl, data=mtcars)
fit4=lm(mpg ~ am + wt + cyl + hp, data=mtcars)
anova(fit1,fit2,fit3,fit4)
```

The final model:
```{r final model}
summary(fit4)
```

This model explains 86.59% of the variance. It may be concluded that on average, manual transmissions have 1.81 more mpg than automatic.


###Appendix

####Appendix 1:
```{r boxplot}
library(ggplot2)
g=ggplot(mtcars, aes(x=am,y=mpg,fill=am)) +
        geom_boxplot(varwidth=T,width=0.5)+
        geom_jitter(shape=16,position=position_jitter(0.1),alpha=0.5)+
        labs(x="Transmission",y="Miles per galon")+
        scale_x_discrete(labels=c("Manual","Automatic"))
g  
```

####Appendix 2:
Residual plots:
```{r residual, fig.height=3.5,fig.width=3.5}
plot(fit4)
```